Minimum Regret Search for Single- and Multi-Task Optimization

### 2016-06-08 2016_ICML/presentation/main.tex
First version of ICML presentation

\begin{frame}{Motivation: Bayesian optimization}

\begin{frame}{Bayesian Optimization}
\textbf{Bayesian optimization}\footcite{shahriari_taking_2016}:
\begin{itemize}
\item black-box optimization problems: $\mathbf{x} = \arg\min_{\mathbf{x} \in \mathcal{X}} f(x)$ of some function $f: \mathcal{X} \to \mathbb{R}$ on some bounded set $\mathcal{X} \subset \mathbb{R}^D$.
\item \emph{probabilistic model} $p(f)$ for $f(\mathbf{x})$, typically a Gaussian process (GP)
\item based on GP posterior of first $n$ observations $\mathcal{D}_n=\{(\mathbf{x}_i, y_i)\}_{i=1}^n$ and an \emph{acquisition function}: decides on a query point $\mathbf{x}_{n+1}$ where $f$ will be evaluated next
\item recommends $\mathbf{\tilde x}_N$ after $N$ queries (optimum of GP or best query point)
\item objective: minimize \emph{simple regret} 
$R_f(\mathbf{\tilde  x}_N) = f(\mathbf{x}^\star) - f(\mathbf{\tilde  x}_N) = \max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{\tilde x}_N)$
\end{itemize}
\end{frame}

\begin{frame}{Entropy Search}
\begin{itemize}
\item Let
   \begin{itemize}
   \item $p^\star(x \vert \mathcal{D}_n)$ denote the posterior distribution (after observing $\mathcal{D}_n$) of the unknown optimizer $\mathbf{x}^\star = \arg\max_{\mathbf{x} \in \mathcal{X}} f(\mathbf{x})$.
   \item $H(\mathbf{x}^\star \vert \mathcal{D}_n)$ denote the differential entropy of $p^\star(x \vert \mathcal{D}_n)$
   \end{itemize}
   \item Entropy Search (ES)\footcite{hennig_entropy_2012} is an information theoretic acquisition fct.:
$$a_{ES}(\mathbf{x}, \mathcal{D}_n) = H(\mathbf{x}^\star \vert \mathcal{D}_n) - \mathbb{E}_{y \vert \mathbf{x}, \mathcal{D}_n}   [H(\mathbf{x}^\star \vert \mathcal{D}_n \cup \{(\mathbf{x}, y))]$$
   \end{itemize}
   \pause

   \vspace*{-.25cm}
   \begin{figure}
   \centering
   \includegraphics[width=0.48\columnwidth]{../pics/regret_illustration_2}
   \visible<3->{
     \includegraphics[width=0.48\columnwidth]{../pics/regret_scatter}
   }
   \vspace*{-.25cm}
   \caption{(Left) GP posterior, probability of maximum $p^\star$, and expected regret ER. (Right) Scatter plot of  position of optimum and regret of $\mathbf{\tilde x} = 1.5$ versus optimum in sample functions.}
   \label{fig:MRS_illustration}
   \end{figure}
\end{frame}

\begin{frame}{Minimum Regret Search}

MRS: acquistion function based directly on external objective of minimizing simple regret
\begin{itemize}
 \item Expected simple regret: $\text{ER}(p)(\mathbf{x}) = \mathbb{E}_{p(f)}[R_f(\mathbf{x})]
= \mathbb{E}_{p(f)}[\max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{
x})]$
 \item For fixed GP $p(f)$, $\arg\min_\mathbf{x} \text{ER}(p)(\mathbf{x})$ corresponds to the maximizer of the GP mean
 \item MRS aims at selecting query points s.t. ER is minimized also with respect to resulting $p(f)$
 \item Myopic choice: select next query point s.t. minimum ER is reduced the most (in expectation)
  $$a_{\text{MRS}^{\text{point}}}(\mathbf{x}^q)
  = \min_{\mathbf{\tilde x}}\text{ER}(p_n)(\mathbf{\tilde x})
   - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[\min_{\mathbf{\tilde x}}  \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]$$
 \item MRS additionally also accounts for uncertainty in minimizer $\mathbf{\tilde x}$ of $\text{ER}$ 
  $$a_{\text{MRS}}(\mathbf{x}^q)
    = \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_n}[\text{ER}(p_n)(\mathbf{\tilde x})] 
     - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[
      \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_{\mathcal{D}_n \cup \{(\mathbf{x}^q, y)\}}}[
       \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]]$$
\end{itemize}

\begin{frame}{Minimum Regret Search Illustration}
\begin{figure}
\centering
\includegraphics[width=0.48\columnwidth]{../pics/acq_comparison}
\end{figure}

\begin{itemize}
\item MRS$^\text{point}$ tends to query in areas which are risky for current recommendation (large simple regret possible)
\item MRS is more smooth than MRS$^\text{point}$ since it accounts for uncertainty in recommendation
\end{itemize}

@@

\begin{frame}{Results: }
\begin{figure}
\centering
\includegraphics[width=.8\textwidth]{../pics/es_analysis}
\label{fig:es_analysis}
\end{figure}

@@

  \vspace*{1.5cm}
  \pause
  \begin{center}
   Thank you for your attention! \\
   Do you have questions, comments, or ideas?
   \end{center}

>>>>

\begin{frame}{Introduction}

\begin{itemize}
  \item Motivation: optimization of expensive target functions
  \item Examples:
  \begin{itemize}
    \item Automated machine learning (computational cost)
    \item Process optimization (economical cost)
    \item Robot learning (supervision cost)
    \item Animal testing (moral cost)
  \end{itemize}
  \pause
  \item Objective: Find approximate optimum of target function with a very small number of function evaluations (``queries'')
  \pause
   \item Approach: Invest significant amounts of computation to determine ``optimal'' sequence of query points
\end{itemize}
\vspace*{4cm}
\end{frame}

\begin{frame}{Bayesian Optimization}
Bayesian optimization\footcite{shahriari_taking_2016}:
\begin{itemize}
\item black-box optimization problems: $\mathbf{x} = \arg\max_{\mathbf{x} \in \mathcal{X}} f(x)$ of some function $f: \mathcal{X} \to \mathbb{R}$ on some bounded set $\mathcal{X} \subset \mathbb{R}^D$.
\item \emph{probabilistic model} $p(f)$ for $f(\mathbf{x})$, typically a Gaussian process (GP)
\item based on GP posterior of first $n$ observations $\mathcal{D}_n=\{(\mathbf{x}_i, y_i)\}_{i=1}^n$ and an \emph{acquisition function}: decides on a query point $\mathbf{x}_{n+1}$ where $f$ will be evaluated next
\item recommends $\mathbf{\tilde x}_N$ after $N$ queries (optimum of GP or best query point)
\item objective: minimize \emph{simple regret} 
$R_f(\mathbf{\tilde  x}_N) = f(\mathbf{x}^\star) - f(\mathbf{\tilde  x}_N) = \max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{\tilde x}_N)$
\end{itemize}
\vspace*{2cm}
\end{frame}

\begin{frame}{Entropy Search}
\begin{itemize}
    \item Let $p^\star(x \vert \mathcal{D}_n)$ denote the posterior distribution (after observing $\mathcal{D}_n$) of the unknown optimizer $\mathbf{x}^\star = \arg\max_{\mathbf{x} \in \mathcal{X}} f(\mathbf{x})$.
    \item and $H(\mathbf{x}^\star \vert \mathcal{D}_n)$ denote the differential entropy of $p^\star(x \vert \mathcal{D}_n)$
    \item Entropy Search (ES)\footcite{hennig_entropy_2012} is an information theoretic acquisition fct.:
$$a_{ES}(\mathbf{x}, \mathcal{D}_n) = H(\mathbf{x}^\star \vert \mathcal{D}_n) - \mathbb{E}_{y \vert \mathbf{x}, \mathcal{D}_n}   [H(\mathbf{x}^\star \vert \mathcal{D}_n \cup \{(\mathbf{x}, y))]$$
   \end{itemize}
   \pause

   \vspace*{-.25cm}
   \begin{figure}
   \centering
   \includegraphics[width=0.48\columnwidth]{../pics/regret_illustration_3}
   \visible<3->{
     \includegraphics[width=0.48\columnwidth]{../pics/regret_scatter}
   }
   \vspace*{-.25cm}
   \caption{\tiny (Left) GP posterior, probability of maximum $p^\star$, and expected regret ER. \visible<3->{(Right) Scatter plot of  position of optimum and regret of $\mathbf{\tilde x} = 1.5$ versus optimum in sample functions.}}
   \label{fig:MRS_illustration}
   \end{figure}
\end{frame}

\begin{frame}{Minimum Regret Search (MRS)}

Acquisition function based directly on minimizing simple regret
\begin{itemize}
 \item Expected simple regret: $\text{ER}(p)(\mathbf{x}) = \mathbb{E}_{p(f)}[R_f(\mathbf{x})]
= \mathbb{E}_{p(f)}[\max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{
x})]$
 \item For fixed GP $p(f)$, $\arg\min_\mathbf{x} \text{ER}(p)(\mathbf{x})$ corresponds to the maximizer of the GP mean
\only<2->{
 \item MRS aims at selecting query points s.t.\ ER is minimized also with respect to resulting $p(f)$
 \item Myopic choice: MRS$^{\text{point}}$ selects next query point s.t.\ minimum ER is reduced the most (in expectation)
  $$a_{\text{MRS}^{\text{point}}}(\mathbf{x}^q)
  = \min_{\mathbf{\tilde x}}\text{ER}(p_n)(\mathbf{\tilde x})
   - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[\min_{\mathbf{\tilde x}}  \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]$$
 \visible<3->{
 \item MRS additionally also accounts for uncertainty in minimizer $\mathbf{\tilde x}$ of $\text{ER}$ 
  $$a_{\text{MRS}}(\mathbf{x}^q)
    = \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_{\mathcal{D}_n}}[\text{ER}(p_n)(\mathbf{\tilde x})] 
     - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[
      \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_{\mathcal{D}_n \cup \{(\mathbf{x}^q, y)\}}}[
       \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]]$$
  }
  }
\end{itemize}
\only<1>{
\begin{center}
\includegraphics[width=0.58\columnwidth]{../pics/regret_illustration_4}
\end{center}
}
\end{frame}

\begin{frame}{Illustration of Acquisition Functions}
\begin{figure}
\centering
\includegraphics[width=0.48\columnwidth]{../pics/acq_comparison_2}
\end{figure}

\vspace*{-.5cm}
\begin{itemize}
\item ES tends to query close to areas where $p^\star$ is large
\item MRS$^\text{point}$ tends to query in areas which are risky for current recommendation (large simple regret possible)
\item MRS is more smooth than MRS$^\text{point}$ since it accounts for uncertainty in recommendation
\end{itemize}

@@

@@

  \vspace*{1.5cm}
  \pause
  \begin{center}
   Thank you for your attention and see you at the poster! \\
   Do you have questions, comments, or ideas?
   \end{center}

\begin{frame}{Results: MRS versus ES}
\begin{figure}
\centering
\includegraphics[width=.8\textwidth]{../pics/es_analysis}
\label{fig:es_analysis}
\end{figure}
